---
title: "Instructors"
format: html
editor: visual
execute:
  message: false
  warning: false
  echo: false
---

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(reactable)
library(ggmap)
library(plotly)
library(ggplot2)
```

```{r}
# Import data and format
instructors <- read_delim("https://raw.githubusercontent.com/Bioconductor/bioconductor-teaching/devel/bioc-instructors.md", delim = '|', trim_ws = TRUE) |> 
  filter(Name != "----------") |> 
  mutate(Status = "Certified")
```

Map and table showing people in the Bioconductor community who are certified Carpentries instructors or in the process of obtaining certification, such as those in the [Bioconductor Carpentries training program](https://bioconductor.github.io/biocblog/posts/2023-02-24-carpentries-update/). If you are certified and don't see your name, please add yourself as described in [this GitHub issue](https://github.com/Bioconductor/bioconductor-teaching/issues/22#issuecomment-1384258079).

See [this page](workshops.html) for Bioconductor Carpentries Workshops.

```{r}
# Get coordinates to map
instructors_geocoded <- instructors |> 
  mutate_geocode(Location)
```

```{r}

# Add instructor trainees from Bioc Carpentries Year 1
trainees_yr1 <- tibble::tribble(
  ~Name, ~Institution, ~Location, ~Status, ~lon, ~lat,
  "Fabricio Almeida-Silva", "VIB-UGent Center for Plant Systems Biology", "Ghent", "Certified - CZI Year 1", 3.7303351, 51.0500182,
  "Anand Ruban Agarvas", "University Hospital Heidelberg", "Heidelberg", "Certified - CZI Year 1", 8.6724335, 49.3987524,
  "Oliver Crook", "University of Oxford", "Oxford", "Certified - CZI Year 1", -1.2577263, 51.7520209,
  "Almut Lütge", "University of Zürich", "Zurich", "Certified - CZI Year 1", 8.541694, 47.3768866,
  "Leo Lahti", "University of Turku", "Turku", "Certified - CZI Year 1", 22.2666303, 60.4518126,
  "Sehyun Oh", "City University of New York", "New York", "Certified - CZI Year 1", -74.0059728, 40.7127753,
  "Umar Ahmad", "Bauchi State University", "Gadau", "In Progress", 10.1661239, 11.8328495,
  "Chia Sin Liew", "University of Nebraska-Lincoln", "Nebraska", "Certified - CZI Year 1", -99.9018131, 41.4925374,
  "Charlotte Hutchings", "University of Cambridge", "Cambridge", "Certified - CZI Year 1", 0.1312729, 52.1950788,
  "Roberto Carlos Álvarez Martínez", "Universidad Autonoma de Queretaro", "Mexico", "In Progress", -102.552784, 23.634501,
  "David Shih", "University of Hong Kong", "Hong Kong", "Certified - CZI Year 1", 114.1693611, 22.3193039,
  "Rodrigo Arcoverde Cerveira da Silva", "Karolinska Institute", "Stockholm", "Certified - CZI Year 1", 18.0685808, 59.3293235,
  "Javier Carpinteyro Ponce", "University of Maryland", "Maryland", "Certified - CZI Year 1", -76.6412712, 39.0457549,
  "Zuguang Gu", "German Cancer Research Center", "Heidelberg", "Certified - CZI Year 1", 8.6724335, 49.3987524,
  "Givanna Putri", "Walter and Eliza Hall Institute of Medical Research", "Melbourne", "Certified - CZI Year 1", 144.9630576, -37.8136276,
  "Zedias Chikwambi", "Chinhoyi University of Technology", "Chinhoyi", "Certified - CZI Year 1", 30.1987245, -17.3622214,
  "Pageneck Chikondowa", "African Institute of Biomedical Science and Technology", "Harare", "Certified - CZI Year 1", 31.0492259, -17.8216288,
  "Marie Hidjo", "African Institute of Biomedical Science and Technology", "Harare", "Certified - CZI Year 1", 31.0492259, -17.8216288
)
```


```{r}
# Add instructor trainees from Bioc Carpentries Year 2
trainees_yr2 <- tibble::tribble(
  ~Name, ~Institution, ~Location, ~Status, ~lon, ~lat,
  "Kevin Stachelek", "University of Southern California", "Los Angeles", "In Progress", -118.285117, 34.0223519,
  "Erick Isaac Navarro Delgado", "University of British Columbia", "Vancouver", "In Progress", -123.1139524, 49.2608724,
  "Pedro Baldoni", "Walter and Eliza Hall Institute of Medical Research", "Melbourne", "Certified - CZI Year 2", 144.9582768, -37.7982827,
  "Princess Rodriguez", "University of Vermont", "Burlington", "In Progress", -73.212072, 44.4758825,
  "Tim Triche", "Van Andel Institute", "Grand Rapids", "Certified - CZI Year 2", -85.6680863, 42.9633604,
  "Izabela Mamede", "Universidade Federal de Minas Gerais", "Belo Horizonte", "Certified - CZI Year 2", -43.9678357, -19.870682,
  "Jacques Serizay", "Institut Pasteur", "Paris", "In Progress", 2.3522219, 48.856614,
  "Robert Ivanek", "University of Basel, Swiss Institute of Bioinformatics", "Basel", "Certified - CZI Year 2", 7.5885761, 47.5595986,
  "Hamdi Furkan Kepenek", "Abdullah Gül University", "Kayseri", "In Progress", 35.4867691, 38.731229,
  "陈秭如", "GuangZhou Lab", "Guangzhou", "In Progress", 113.264435, 23.129163,
  "Amal Boukteb", "National Institute for Agricultural Research of Tunisia (INRAT)", "Tunis", "Certified - CZI Year 2", 10.1815316, 36.8064948,
  "Arnab Mukherjee", "Manipal Institute of Technology", "Manipal", "Certified - CZI Year 2", 74.7928923, 13.3532052,
  "Harvinder Singh", "Department of Translational and Regenerative Medicine, PGIMER", "Chandigarh", "In Progress", 76.779418, 30.733315,
  "Janani Ravi", "University of Colorado Anschutz Medical Campus", "Aurora", "In Progress", -104.991531, 39.742043
)
```


```{r}
# Combine all data
all <- bind_rows(instructors_geocoded, trainees_yr1, trainees_yr2) |> 
  mutate(Status = factor(Status, levels = c(
  "Certified", 
  "Certified - CZI Year 1", 
  "Certified - CZI Year 2", 
  "In Progress")
  ))
```

```{r}
# Map of instructors
world_map <- map_data("world") |> 
  filter(region != "Antarctica")

gg <- ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "grey85", color = "white") +  # Pale grey countries with white borders
  geom_jitter(data = all, aes(x = lon, y = lat, text = paste(Name, Institution, sep = "\n"), 
                              colour = Status, fill = Status, shape = Status, size = Status)) +
  scale_colour_manual(values = c(
    "Certified" = "#0072B2", 
    "Certified - CZI Year 1" = "#009E73", 
    "Certified - CZI Year 2" = "#CC79A7", 
    "In Progress" = "#E69F00")) +
  scale_fill_manual(values = c(
    "Certified" = "#0072B2", 
    "Certified - CZI Year 1" = "#009E73", 
    "Certified - CZI Year 2" = "#CC79A7", 
    "In Progress" = "#E69F00")) +
  scale_shape_manual(values = c(
    "Certified" = 21,  # Filled Circle
    "Certified - CZI Year 1" = 22,  # Filled Square
    "Certified - CZI Year 2" = 24,  # Filled Diamond
    "In Progress" = 21  # Filled Circle
  )) +
  scale_size_manual(values = c(
    "Certified" = 1, 
    "Certified - CZI Year 1" = 1,  
    "Certified - CZI Year 2" = 1,  
    "In Progress" = 1
  )) +
  coord_sf(xlim = c(-180, 180), ylim = c(-60, 90), expand = FALSE) +  # Adjust x and y limits to minimize white space
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    axis.line = element_blank(),  # No axis lines
    axis.text = element_blank(),  # No axis text
    axis.ticks = element_blank(),  # No axis ticks
    plot.background = element_rect(fill = "white", color = NA)  # Ensure no background around the map
  )

ggplotly(gg, tooltip = "text") |> 
  layout(xaxis = list(showline = FALSE, showticklabels = FALSE, zeroline = FALSE, title_standoff = 5),
         yaxis = list(showline = FALSE, showticklabels = FALSE, zeroline = FALSE))

```

<br></br>

```{r}
# Table of instructors
all |>   
    select(Name, Institution, Location, Status) |> 
  reactable(columns = list(
    Name = colDef(cell = function(value, index) {
      if (is.na(all$www[index])) {
        # If URL is NA, just display the name without a link
        value
        } else {
          # Render as a proper HTML link
          htmltools::tags$a(href = all$www[index], target = "_blank", value)
          }
      })
    ), 
    defaultSorted = "Name",
    defaultPageSize = 100)

```

However, looking at the world population density below from [Our World in Data](https://ourworldindata.org/grapher/population-density), we can see there are highly dense areas that have no or few trainers (e.g. India, China, Middle East, Africa, Latin America) so we still have a bit of a way to go! 🌍🌎🌏

<iframe src="https://ourworldindata.org/grapher/population-density" loading="lazy" style="width: 100%; height: 600px; border: 0px none;">

</iframe>
