---
title: "Instructors"
format: html
editor: visual
execute:
  message: false
  warning: false
  echo: false
---

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(reactable)
library(ggmap)
library(plotly)
```

```{r}
# import data and format
instructors <- read_delim("https://raw.githubusercontent.com/Bioconductor/bioconductor-teaching/devel/bioc-instructors.md", delim = '|', trim_ws = TRUE) |> 
  filter(Name != "----------") |> 
  mutate(Status = "Certified")
```

Map and table showing people in the Bioconductor community who are certified Carpentries instructors or in the process of obtaining certification, such as those in the [Bioconductor Carpentries training program](https://bioconductor.github.io/biocblog/posts/2023-02-24-carpentries-update/). If you are certified and don't see your name, please add yourself as described in [this GitHub issue](https://github.com/Bioconductor/bioconductor-teaching/issues/22#issuecomment-1384258079).

See [this page](workshops.html) for Bioconductor Carpentries Workshops.

```{r}
# get coordinates to map
instructors_geocoded <- instructors |> 
  mutate_geocode(Location)
```

```{r}
#| echo: false

# add instructor trainees from Bioc Carpentries 
trainees <- tibble::tribble(
                                  ~Name,                                                   ~Institution,    ~Location,  ~Status,        ~lon,        ~lat,
               "Fabricio Almeida-Silva",                   "VIB-UGent Center for Plant Systems Biology",      "Ghent",  "Newly Certified",  3.7303351,  51.0500182,
                  "Anand Ruban Agarvas",                               "University Hospital Heidelberg", "Heidelberg",  "In Progress",     8.6724335,  49.3987524,
                         "Oliver Crook",                                          "University of Oxford",     "Oxford",  "In Progress",    -1.2577263,  51.7520209,
                          "Almut Lütge", "University of Zürich",     "Zurich",  "Newly Certified",      8.541694,  47.3768866,
                            "Leo Lahti",                                          "University of Turku",      "Turku",  "Newly Certified",    22.2666303,  60.4518126,
                            "Sehyun Oh",                                  "City University of New York",   "New York",  "Newly Certified",   -74.0059728,  40.7127753,
                           "Umar Ahmad",                                      "Bauchi State University",      "Gadau",  "In Progress",    10.1661239,  11.8328495,
                        "Chia Sin Liew",                               "University of Nebraska-Lincoln",   "Nebraska",  "Newly Certified",   -99.9018131,  41.4925374,
                  "Charlotte Hutchings",                                      "University of Cambridge",  "Cambridge",  "Newly Certified",     0.1312729,  52.1950788,
                   "Roberto Carlos Álvarez Martínez",                       "Universidad Autonoma de Queretaro",  "Mexico",  "In Progress",    -102.552784,  23.634501,
                           "David Shih",                                      "University of Hong Kong",  "Hong Kong",  "Newly Certified",   114.1693611,  22.3193039,
  "Rodrigo Arcoverde Cerveira da Silva",                                        "Karolinska Institute",  "Stockholm",  "Newly Certified",    18.0685808,  59.3293235,
             "Javier Carpinteyro Ponce",                                       "University of Maryland",   "Maryland",  "Newly Certified",   -76.6412712,  39.0457549,
                           "Zuguang Gu",                                "German Cancer Research Center", "Heidelberg",  "Newly Certified",     8.6724335,  49.3987524,
                        "Givanna Putri",          "Walter and Eliza Hall Institute of Medical Research",  "Melbourne",  "Newly Certified",   144.9630576, -37.8136276,
                     "Zedias Chikwambi",                            "Chinhoyi University of Technology",   "Chinhoyi",  "Newly Certified",    30.1987245, -17.3622214,
                  "Pageneck Chikondowa",       "African Institute of Biomedical Science and Technology",     "Harare",  "Newly Certified",    31.0492259, -17.8216288,
                          "Marie Hidjo",       "African Institute of Biomedical Science and Technology",     "Harare",  "Newly Certified",    31.0492259, -17.8216288
  )
```

```{r}
# Combine
all <- bind_rows(instructors_geocoded, trainees) |> 
  mutate(Status = factor(Status, levels = c("Certified", "Newly Certified", "In Progress")))
```

```{r}
# Map of instructors
world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")

gg <- ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "grey") +
  geom_jitter(data = all, aes(x = lon, y = lat, text = paste(Name, Institution, sep = "\n"), colour = Status, shape = Status, size = Status)) +
  scale_colour_manual(values = c("Certified"="blue", "Newly Certified"="gold", "In Progress"="red")) +
  scale_shape_manual(values = c("Certified"=20, "Newly Certified"=18, "In Progress"=20)) +
  scale_size_manual(values = c("Certified"=1, "Newly Certified"=1, "In Progress"=1)) +
  theme_void()

ggplotly(gg, tooltip = "text") |> 
  layout(width=700, height=450,
         legend = list(orientation = "h", x = 0.25, y = -0.02))
```

<br></br>

```{r}
# Table of instructors
all |>   
    select(Name, Institution, Location, Status) |> 
  reactable(columns = list(
    Name = colDef(cell = function(value, index) {
      # Render as a link
      htmltools::tags$a(href = all$www[index], target = "_blank", as.character(value))
      })
    ), 
    defaultSorted = "Name",
    defaultPageSize = 100)

```

However, looking at the world population density below from [Our World in Data](https://ourworldindata.org/grapher/population-density), we can see there are highly dense areas that have no or few trainers (e.g. India, China, Middle East, Africa, Latin America) so we still have a bit of a way to go! 🌍🌎🌏

<iframe src="https://ourworldindata.org/grapher/population-density" loading="lazy" style="width: 100%; height: 600px; border: 0px none;">

</iframe>
